---
title: "Assignment 2"
author: "Annie"
date: "24/09/2019"
output: github_document
always_allow_html: yes
---

```{r setup, warning = FALSE, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(gapminder)
library(tidyverse)
```

1.1. This table shows gapminder data for Canada, USA, and Mexico in the 1970's.

```{r}
gapminder_70 <- gapminder %>%
  filter(country == 'Canada' | country == 'United States' | country == 'Mexico') %>% 
  filter(year >= 1970 & year < 1980)
gapminder_70
```

1.2. To hone in on my dataset from 1.1, I used the pipe operator %>% to select “country” and “gdpPercap”.

```{r}
gapminder_select <- gapminder_70 %>%
  select(country, gdpPercap)
gapminder_select
```

1.3. This table shows all entries in the gapminder data that have experienced a drop in life expectancy.

```{r}
lifeexp_drop <- gapminder %>%
  arrange(country, year) %>%
  group_by(country) %>%
  mutate(lifeExplag = lag(lifeExp)) %>%
  mutate(lifeExpdif = lifeExp - lifeExplag) %>%
  filter(lifeExpdif < 0)
lifeexp_drop
```

1.4. Gapminder data is filtered to show the max GDP per capita experienced by each country.

```{r}
max_gdp <- gapminder %>%
  group_by(country) %>%
  summarise(max(gdpPercap))
max_gdp
```

1.5. This scatterplot shows Canada’s life expectancy vs. GDP per capita.

```{r}
gapminder %>% 
  filter(country == 'Canada') %>%
  select(lifeExp, gdpPercap) %>%
    ggplot(aes(lifeExp, gdpPercap)) +
    geom_point() +
    xlab("Life expectancy") +
    scale_y_log10("GDP per capita", labels = scales::dollar_format())
```

2. Below I explore these two variables: 

Quantitative variable: Life Expectancy
Categorical variable: Continent

**Life Expectancy**
Below is summary statistics for the variable Life Expectancy within the gapminder dataset.
```{r}
summary(gapminder$lifeExp)
```

Below is the mean Life Expectancy by continent.
```{r}
cont_life <- gapminder %>%
  group_by(continent) %>%
  summarise(mean(lifeExp))
cont_life
```

Below is a histogram of Life Expectancy within the gapminder dataset. The dashed line represents the mean Life Expectancy.
```{r}
gapminder %>%
  mutate(lifeExp_round = round(lifeExp, 0)) %>%
    ggplot(aes(x = lifeExp_round)) +
    geom_histogram(binwidth = 1, color="black", fill="lightblue") +
    xlab("Life expectancy") +
    geom_vline(aes(xintercept = mean(lifeExp)), 
               color="darkblue", linetype="dashed", size=1)
```

**Continent**
Below is the number of countries within each continent.
```{r}
uniquecontinent <- unique(gapminder[1:2]) 

table(uniquecontinent$continent)
```

Below is a table of countries associated with each continent.
```{r}
uniquecontinent %>%
  arrange(continent)
```

3. It is important to understand trends in GDP per Capita. The first graph is a scatterplot of the mean GDP per Capita by continent across time. 

```{r}
gdpmean <- gapminder %>%
  group_by(continent, year) %>%
  summarise(mean(gdpPercap))

gdpmean %>%
    ggplot(aes(year, `mean(gdpPercap)`, color = continent)) +
    geom_point() +
    xlab("Year") +
    scale_y_log10('GDP per Capita', labels = scales::dollar_format()) +
    geom_smooth(method = lm, se = FALSE, size = 0.5)
```

The second graph is a histogram of GDP per Capita across all countries in 2007. 
```{r}
gapminder %>%
  filter(year == 2007) %>%
  group_by(continent) %>%
    ggplot(aes(x = gdpPercap, color = continent)) +
    geom_histogram(binwidth = 5000, color="black", fill="lightblue") +
    xlab("GDP per Capita") +
    geom_vline(aes(xintercept = mean(gdpPercap)), 
               color="darkblue", linetype="dashed", size=1)
```

Extra Part 1

*Evaluate this code and describe the result. Presumably the analyst’s intent was to get the data for Rwanda and Afghanistan. Did they succeed? Why or why not? If not, what is the correct way to do this?*

```{r}
filter(gapminder, country == c("Rwanda", "Afghanistan"))
```

The code does filter gapminder's data to Rwanda and Afghanistan, but it is not complete. It is missing data from certain years for both Rwanda and Afghanistan. Instead, the code should be:

```{r}
gapminder_AfRw <- gapminder %>%
  filter(country == 'Afghanistan' | country == 'Rwanda')
gapminder_AfRw
```

