---
title: "Assignment 1: Gapminder"
output: github_document
always_allow_html: yes
---

```{r setup, echo=FALSE}
library(gapminder)
library(tibble)
library (DT)
```

## Gapminder
We are going to explore Gapminder, a dataset.
What information is present in the dataset? Below are Gapminder's column headings:
```{r gapminder}
columns <- names(gapminder)
columns
```

## Smallest population Part 1
Which country in the world has the smallest population size?
```{r, echo=FALSE}
min_pop <- subset(gapminder, pop==min(pop), select = country:pop)
min_pop
```

## Smallest population Part 2
The smallest population size in Gapminder is from a country in 1952. 
Does the same country currently have the smallest population size in the world?
*Current = 2007, the most recent year available in Gapminder*
```{r, echo=FALSE}
recent <- subset(gapminder, year=='2007')
min_pop_recent <- subset(recent, pop==min(pop), select = country:pop)
min_pop_recent
```

## Population range
Globally, what is the range in population sizes (circa 2007)?
```{r, echo=FALSE}
recent <- subset(gapminder, year=='2007')
range_pop <- (range(recent$pop))
range_pop
```