---
title: "Homework 04"
author: "Annie"
date: "08/10/2019"
output: github_document
---

This assignment explores tidy data using the dataset gapminder.
```{r setup, include=FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(gapminder)
library(tidyverse)
library(ggplot2)
guest <- read_csv("https://raw.githubusercontent.com/STAT545-UBC/Classroom/master/data/wedding/attend.csv")
email <- read_csv("https://raw.githubusercontent.com/STAT545-UBC/Classroom/master/data/wedding/emails.csv")
```

##Univariate Option 1
Below I have made a tibble with one row per year, and columns for life expectancy for the Germany and China.
```{r}
gapminder_JaCh <- gapminder %>%
  filter(country == 'Japan' | country == 'China')
gapminder_JaCh

JaCh_wide <- gapminder_GeCh %>%
  pivot_wider(id_cols = year, 
              names_from = country, 
              values_from = lifeExp)
JaCh_wide
```

Below is a scatterplot of the life expectancy of Russia vs. China.

```{r}
JaCh_wide %>%
  ggplot(aes(year))+
    geom_point(aes(y = Japan))+
    geom_point(aes(y = China)) +
    labs(x = "Year", y = "Life Expectancy") +
    scale_color_discrete("Legend")
```


Now, I will relengthen the dataset. 

```{r}
JaCh_long <- JaCh_wide %>%
  pivot_longer(cols = -year,
               names_to = "Country",
               values_to = "Life Expectancy")
#For cols, I could also insert Japan:China
JaCh_long
```

##Multivariate Option 2

Below I will compute the mean life expectancy and GDP per capita for all possible combinations of continent and year.

```{r}
meanlife_gdp <- gapminder %>%
  group_by(continent, year) %>%
  summarize(meanlife = mean(lifeExp), meangdp = mean(gdpPercap))

meanlifegdp_wide <- meanlife_gdp %>%
  pivot_wider(id_cols = c(year),
              names_from = continent,
              names_sep = "_",
              values_from = c(meanlife, meangdp))
meanlifegdp_wide
```

Next, I will lengthen the data.

```{r}
meanlifegdp_long <- meanlifegdp_wide %>%
  pivot_longer(cols = contains("mean"), 
               names_to = c("continent", ".value"),
               names_sep = "_")
meanlifegdp_long
```

##Table Joins

For this exercise, I am going to add the appropriate email address for every guest. 

```{r}

```

